---
title: "Spatial Analysis Figures"
subtitle: "Area 25 Chum Salmon"
author: "Coastland"
date: "`r Sys.Date()`"
papersize: ledger
geometry: margin = 1.75cm
toc: TRUE
output:
  pdf_document: default
  pescdata.filt_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE,message=FALSE,warning=FALSE}
#load libraries
knitr::opts_chunk$set(echo = TRUE, fig.pos = "!h")
options(knitr.table.format ='pdf' ) 

library(data.table)
library(tidyverse)
library(corrplot)
library(ggpubr)
library(zoo)
library(ggdendro)
library(dendextend)
library(kableExtra)
library(tinytex)

options(scipen=10000)

rps.clean <- readRDS("RDS/a25cm_rps.clean.rds")

```


```{r functions, echo=FALSE,message=FALSE,warning=FALSE}
#function to extract data and group it

getridofggplotlinesbetweenpoints<-function(data,colstokeep,yearcolumnname) {
  d2<-data%>%select(year=yearcolumnname,colstokeep)
  idx <- c(1, diff(d2$year))
  i2 <- c(1,which(idx != 1), nrow(d2)+1)
  d2$grp <- rep(1:length(diff(i2)), diff(i2))
  d2
}
```

\newpage

```{r load escapement data,echo=FALSE}

escdata.raw<-fread("data/area 25 escapement data all.csv")
escdata.filt<-fread("data/area 25 escapement data filtered.csv")%>%select(-V1)

```

# Summary statistics 

## Bubbleplot of escapement by enhancement rank

```{r escapement bubble plot, fig.align="center",fig.height=8,fig.width=6,fig.cap="Escapement to area streams by enhancement rank.",echo=FALSE,warning=FALSE,message=FALSE}

par(mfrow = c(1, 1))
#bubbleplot of escapements for whole area
ggplot(escdata.raw,aes(x=Year,y=SYS_NM,color=Enhancement_Rank,size=escapement))+
  geom_point(alpha=.5)+
  scale_colour_manual(breaks=c("NONE","LOW","HIGH"),
                      values=c("grey50","steelblue","darkred"))+
  labs(y="Escapement",color="Enhancement",x="Year",size="Escapement")+
  theme_bw()+
  theme(legend.position="bottom")+
  guides(colour = guide_legend(nrow = 4))+
  guides(size = guide_legend(nrow = 4))+ 
  theme(axis.text.x = element_text(size = 7))

```

\newpage

```{r escapement filtered streams only, fig.align="center",fig.height=8,fig.width=7,fig.cap="Escapement to area streams by enhancement rank.",echo=FALSE,warning=FALSE,message=FALSE}

ggplot(escdata.filt,aes(x=Year,y=escapement,color=Enhancement_Rank))+
  geom_point()+geom_line()+theme_bw()+
  facet_wrap(~SYS_NM,ncol=3,scale="free_y")+
  labs(y="Escapement",color="Enhancement Level")+
  theme(legend.position="bottom")

```


\newpage

## Releases by system

```{r, echo=FALSE,message=FALSE,warning=FALSE,fig.height=7,fig.width=7,fig.cap="Chum releases to Area 25 by release site and release stage."}

swvi.chum<-fread("data/swvi chum releases filtered.csv")

ggdata.rel <- swvi.chum
ggdata.rel <- ggdata.rel[,2:14]

idx <- c(1, diff(ggdata.rel$RELEASE_YEAR))
i2 <- c(1,which(idx != 1), nrow(ggdata.rel)+1)
ggdata.rel$grp <- rep(1:length(diff(i2)), diff(i2))

ggplot(ggdata.rel,aes(x=RELEASE_YEAR,y=releases,color=RELEASE_STAGE_NAME,group=grp))+
  geom_line()+
  geom_point(size = 0.7)+
  scale_color_brewer(palette="Set1")+
  facet_wrap(~`site-stock`,ncol=3)+
  theme_bw()+
  labs(x="Release Year",y="Releases",color="Release Stage")+
  theme(legend.position = "bottom")

```

\newpage

## Table of A25 streams by inlet and distance from Conuma enhancement


```{r, Kable of distances by inlet, echo = FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.height = 7, fig.width = 7}

streaminlet <- fread("data/CM A25 filtered stream list_with inlets.csv")

colnames(streaminlet) <- c("System name", "NOTNAS", "Inlet", "Subinlet", "Dist. from enhancement")
colaligns <- c("l", "c", "l", "l", "r")
streaminlet$`System name` <- str_to_title(streaminlet$`System name`)
streaminlet$Inlet <- str_to_sentence(streaminlet$Inlet)
streaminlet$Subinlet <- str_to_sentence(streaminlet$Subinlet)

streaminlet %>%
  kbl(align = colaligns, format.args = list(big.mark = ",")) %>%
  kable_paper(full_width = F, html_font = "Courier New") %>%
  kable_styling(latex_options = "HOLD_position")

```

\newpage

## Hatchery releases to area
```{r releases for area 25, fig.align="center", fig.height=5, fig.width=8, fig.cap="Total hatchery chum salmon releases in Area 25",echo=FALSE,warning=FALSE,message=FALSE}

releases.raw <- read_csv("data/Area 25 chum releases totals.csv")

ggplot(releases.raw, aes(x = RELEASE_YEAR, y = totreleases/1000000)) + 
  geom_point() +
  geom_line() + 
  geom_smooth() + 
  xlab("Year") + 
  ylab("Total A25 chum salmon releases (millions)") +
  theme_bw() +
  scale_fill_grey()

```

\newpage

## Escapement by enhancement rank per system

```{r metric plots,fig.align="center",fig.height=8,fig.width=6,fig.cap="Various plots for escapement and transformations.",echo=FALSE,warning=FALSE,message=FALSE}
#plot streams by enhancement

p1<-ggplot(escdata.filt,aes(x=Year,y=escapement,color=Enhancement_Rank,group=SYS_NM))+
  geom_line()+
  theme_bw()+
  labs(y="Escapement",title="Raw Escapement",color="Enhancement Rank")

p2<-ggplot(escdata.filt,aes(x=Year,y=esc.log,color=Enhancement_Rank,group=SYS_NM))+
 geom_line()+
  theme_bw()+
  labs(y="log(Escapement)",title="Logged Escapement")

p3<-ggplot(escdata.filt,aes(x=Year,y=z,color=Enhancement_Rank,group=SYS_NM))+
  geom_line()+
  geom_hline(yintercept=0)+
  theme_bw()+
  labs(y="Z-scores (Escapement)",title="Z-scores Escapement")

p4<-ggplot(escdata.filt,aes(x=Year,y=esc.stand,color=Enhancement_Rank,group=SYS_NM))+
  geom_line()+
  theme_bw()+
  labs(y="Standardized Escapement (pavg)",title="Escapement/Average (Pavg)")

ggarrange(p1,p2,p3,p4,ncol=1,common.legend = TRUE,align='v',legend="bottom")

```

\newpage

## Moving average and LOESSS fit on enhancement ranking of log escapements

```{r moving average and LOESS plots, echo=FALSE,warning=FALSE,message=FALSE,fig.align="center",fig.height=5,fig.width=6,fig.cap="Moving average and LOESS fits on logged escapement by enhancement ranking."}

t1<-ggplot(escdata.filt,aes(x=Year,y=m.ave,color=Enhancement_Rank,group=SYS_NM))+
  geom_line()+
  theme_bw()+
  labs(y="Moving Average",title="Moving-Average",color="Enhancement Rank")

t2<-ggplot(escdata.filt,aes(x=Year,y=z,color=Enhancement_Rank,group=SYS_NM))+
  geom_smooth(method="loess",se=FALSE,span=.1,linewidth=.7)+
  geom_hline(yintercept=0)+
  theme_bw()+
  labs(y="Z-score",title="Loess fit on Z-scores")

ggarrange(t1,t2,ncol=1,common.legend = TRUE,legend = "bottom",align='v')

```

\newpage

## Z-scores pre- and post-enhancement

```{r prepost boxplot,echo=FALSE,warning=FALSE,message=FALSE,fig.align="center",fig.height=7,fig.width=6,fig.cap="Z-scores pre- and post-enhancement."}
ggplot(escdata.filt,aes(y=SYS_NM,x=z,fill=prepost))+
  geom_boxplot()+
  geom_vline(xintercept=0)+
  theme_bw()+
  labs(y="System Name",x="Z-score")
```

\newpage

## Log RPS by stream

```{r rps figure pre post with lm, echo=FALSE, warning=FALSE, message=FALSE ,fig.align="center", fig.height=7, fig.width=6 ,fig.cap="Log RPS by stream with linear regression fits pre- and post-first year of enhancement."}

a25.rps<-fread("data/area 25 chum rps.csv")

gga25.logrps<-a25.rps%>%select(Year,prepost,logrps,SYS_NM)%>%arrange(SYS_NM,prepost,Year)

idx <- c(1, diff(gga25.logrps$Year))
i2 <- c(1,which(idx != 1), nrow(gga25.logrps)+1)
gga25.logrps$grp <- rep(1:length(diff(i2)), diff(i2))

ggplot(gga25.logrps,aes(x=Year,y=logrps,color=prepost,group=grp))+
  geom_line(alpha=.5)+geom_point(alpha=.5)+
  scale_color_brewer(palette="Set1")+
  geom_smooth(method='lm',se=TRUE)+
  facet_wrap(~SYS_NM,ncol=3,scale="free_y")+
  theme_bw()+
  geom_hline(yintercept=0,linetype="dashed")+
  labs(x="Brood Year",y="log(Recruits per spawner)",color="Period")+
  theme(legend.position="bottom")
```

\newpage

## Log RPS by stream boxplot

```{r rps figure boxplot,echo=FALSE,warning=FALSE,message=FALSE,fig.align="center",fig.height=7,fig.width=6,fig.cap="Log RPS by stream boxplot for pre- and post-start of enhancement period."}

ggplot(gga25.logrps,aes(y=SYS_NM,x=logrps,fill=prepost))+
  geom_boxplot(notch=F)+
  scale_fill_brewer(palette="Set1")+
  geom_vline(xintercept=0)+
  theme_bw()+
  labs(y="System Name",x="log(Recruits per spawner)",title="Area 25 Chum")+
  theme(legend.position="bottom")

```

```{r correlations for all metrics, echo = FALSE, warning = FALSE, message = FALSE}

#prep data for correlations

# Merge stream inlet table into escdata.filt

streaminlet <- fread("data/CM A25 filtered stream list_with inlets.csv")

rps.clean <- merge(streaminlet, rps.clean, by = "SYS_NM", all = TRUE)

escdata.filt<-merge(escdata.filt,streaminlet,by="SYS_NM")

#old way
#escdata.filt <- escdata.filt %>%
#  inner_join(streaminlet, escdata.filt, by = "SYS_NM")

# Factor out inlet and create escdata.filt of inlet colours

escdata.filt$inlet <- as.factor(escdata.filt$inlet)
inlets <- as.data.frame(levels(escdata.filt$inlet))
inlets$colour <- c("red", "blue", "darkgreen")
colnames(inlets) <- c("inlet", "colour")


# Correlation analysis by z-score

cor.data.z<-escdata.filt%>%select(SYS_NM,Year,z)%>%
  pivot_wider(names_from="Year",values_from="z")%>%
  data.frame()

rownames(cor.data.z)<-cor.data.z$SYS_NM

cor.data.z<-cor.data.z%>%select(-SYS_NM)

cor.z<-cor(t(cor.data.z),use="pairwise.complete.obs")

# Dendrogram, factored by inlet, by z-score

dist.cor.z<- dist(cor.z)
clust.dist.cor.z <- hclust(dist.cor.z)

dendclust.z <- dendro_data(clust.dist.cor.z)
zlabs <- as.data.frame(dendclust.z[["labels"]][["label"]])

zlabs$order <- 1:length(zlabs[,1])
colnames(zlabs) <- c("SYS_NM", "order")
zlabs <- merge(streaminlet, zlabs, by = "SYS_NM")
zlabs <- merge(inlets, zlabs, by = "inlet")
zlabs <- zlabs %>%
  arrange(order)
zlabcolours <- zlabs$colour
zlabnames <- zlabs$SYS_NM
inletlist <- as.factor(zlabs$inlet) 
  
dendclust.z[["labels"]][["inlet"]] <- inletlist

nameslookup <- as.data.frame(c("Tlupana", "Canton", "Conuma", "Leiner", "Deserted", "Sucwoa", 
                 "Mooyah", "Ransom", "Burman", "Kleeptee", "Espinosa",
                 "L. Zeballos", "Hoiss", "Lord", "Zeballos", "Tahsis", "Tsowwin",
                 "Hammond", "Marvinas", "Brodick", "Owossitsa", "Chum", 
                 "Mamat", "Park"))

colnames(nameslookup) <- "shortnames"

nameslookup$SYS_NM <- zlabnames  

nameslookup$erank <- c("red", "red", "red", "black", "#3399FF", "red",
                       "black", "black", "black", "black", "black",
                       "black", "black", "black", "red", "black", "black",
                       "black", "black", "black", "black", "black",
                       "black", "black")
  
dendclust.z[["labels"]][["inlet"]] <- inletlist

xlenz <- .95 * length(zlabs$SYS_NM)
ylenz <- .95 * (max(dendclust.z[["segments"]][["y"]]))

 

# Correlation analysis by log escapement

cor.data.e<-escdata.filt%>%select(SYS_NM,Year,esc.log)%>%pivot_wider(names_from="Year",values_from="esc.log")%>%data.frame()

rownames(cor.data.e)<-cor.data.e$SYS_NM

cor.data.e<-cor.data.e%>%select(-SYS_NM)

cor.e<-cor(t(cor.data.e),use="pairwise.complete.obs")

# Dendrogram, factored by inlet, by log escapement

dist.cor.e<- dist(cor.e)
clust.dist.cor.e <- hclust(dist.cor.e)

dendclust.e <- dendro_data(clust.dist.cor.e)
elabs <- as.data.frame(dendclust.e[["labels"]][["label"]])
elabs$order <- 1:length(elabs[,1])
colnames(elabs) <- c("SYS_NM", "order")
elabs <- merge(streaminlet, elabs, by = "SYS_NM")
elabs <- merge(inlets, elabs, by = "inlet")
elabs <- elabs %>%
  arrange(order)
elabcolours <- elabs$colour
elabnames <- elabs$SYS_NM
inletlist <- as.factor(elabs$inlet)
  
dendclust.e[["labels"]][["inlet"]] <- inletlist

xlene <- .95 * length(elabs$SYS_NM)
ylene <- .95 * (max(dendclust.e[["segments"]][["y"]]))


# Correlation analysis by standardized escapement
cor.data.s<-escdata.filt%>%select(SYS_NM,Year,esc.stand)%>%pivot_wider(names_from="Year",values_from="esc.stand")%>%data.frame()

rownames(cor.data.s)<-cor.data.s$SYS_NM

cor.data.s<-cor.data.s%>%select(-SYS_NM)

cor.s<-cor(t(cor.data.s),use="pairwise.complete.obs")

# Dendrogram, factored by inlet, by standardized escapement

dist.cor.s<- dist(cor.s)
clust.dist.cor.s <- hclust(dist.cor.s)

dendclust.s <- dendro_data(clust.dist.cor.s)
slabs <- as.data.frame(dendclust.s[["labels"]][["label"]])
slabs$order <- 1:length(slabs[,1])
colnames(slabs) <- c("SYS_NM", "order")
slabs <- merge(streaminlet, slabs, by = "SYS_NM")
slabs <- merge(inlets, slabs, by = "inlet")
slabs <- slabs %>%
  arrange(order)
slabcolours <- slabs$colour
slabnames <- slabs$SYS_NM
inletlist <- as.factor(slabs$inlet)
  
dendclust.s[["labels"]][["inlet"]] <- inletlist

xlens <- .95 * length(slabs$SYS_NM)
ylens <- .95 * (max(dendclust.s[["segments"]][["y"]]))

# Correlation analysis by moving average
cor.data.m<-escdata.filt%>%select(SYS_NM,Year,m.ave)%>%pivot_wider(names_from="Year",values_from="m.ave")%>%data.frame()

rownames(cor.data.m)<-cor.data.m$SYS_NM

cor.data.m<-cor.data.m%>%select(-SYS_NM)

cor.m<-cor(t(cor.data.m),use="pairwise.complete.obs")

# Dendrogram, factored by inlet, by moving average

dist.cor.m<- dist(cor.m)
clust.dist.cor.m <- hclust(dist.cor.m)
dendclust.m <- dendro_data(clust.dist.cor.m)

mlabs <- as.data.frame(dendclust.m[["labels"]][["label"]])
mlabs$order <- 1:length(mlabs[,1])
colnames(mlabs) <- c("SYS_NM", "order")
mlabs <- merge(streaminlet, mlabs, by = "SYS_NM")
mlabs <- merge(inlets, mlabs, by = "inlet")
mlabs <- mlabs %>%
  arrange(order)
mlabcolours <- mlabs$colour
mlabnames <- mlabs$SYS_NM
inletlist <- as.factor(mlabs$inlet)
  
dendclust.m[["labels"]][["inlet"]] <- inletlist

xlenm <- .95 * length(mlabs$SYS_NM)
ylenm <- .95 * (max(dendclust.m[["segments"]][["y"]]))

cor.data.rps<-rps.clean%>%select(SYS_NM,Year,rps)%>%
  pivot_wider(names_from="Year",values_from="rps")%>%
  data.frame()
rownames(cor.data.rps)<-cor.data.rps$SYS_NM
cor.data.rps<-cor.data.rps%>%select(-SYS_NM)
cor.rps<-cor(t(cor.data.rps),use="pairwise.complete.obs")
dist.cor.rps<- dist(cor.rps)
clust.dist.cor.rps <- hclust(dist.cor.rps)
dendclust.rps <- dendro_data(clust.dist.cor.rps)

cor.data.rpslog<-rps.clean%>%select(SYS_NM,Year,logrps)%>%
  pivot_wider(names_from="Year",values_from="logrps")%>%
  data.frame()
rownames(cor.data.rpslog)<-cor.data.rpslog$SYS_NM
cor.data.rpslog<-cor.data.rpslog%>%select(-SYS_NM)
cor.rpslog<-cor(t(cor.data.rpslog),use="pairwise.complete.obs")
dist.cor.rpslog<- dist(cor.rpslog)
clust.dist.cor.rpslog <- hclust(dist.cor.rpslog)
dendclust.rpslog <- dendro_data(clust.dist.cor.rpslog)


```

\newpage

# Correlation analyses

## Correlation plots by metric

```{r correlation plots, echo=FALSE,warning=FALSE,message=FALSE,fig.align="center",fig.height=9.5,fig.width=8.5,fig.cap="Cross correlation plots to compare metrics."}

# Corplots to knit

#png(file = "figures/corrplots.png", width = 2000, height = 3000, pointsize = 48)

par(oma=c(0,0,0,0))
par(mfrow=c(3,2))

corrplot(cor.z, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.45, title = "Z-scores", mar=c(0,0,3,0))
corrplot(cor.e, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.45, title = "Log escapement", mar=c(0,0,3,0))
corrplot(cor.s, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.45, title = "Standardized escapement", mar=c(0,0,3,0))
corrplot(cor.m, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.45, title = "Moving average", mar=c(0,0,3,0))
corrplot(cor.rps, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.45, title = "RPS", mar=c(0,0,3,0))
corrplot(cor.rpslog, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.45, title = "Log RPS", mar=c(0,0,3,0))

#dev.off()
```

\newpage

```{r, Make colour vectors for dendrogram inlets, message=FALSE, warning=FALSE, echo=FALSE, include = FALSE}

zlabs <- as.data.frame(dendclust.z[["labels"]][["label"]])
colnames(zlabs) <- "SYS_NM"
zlabs$order <- c(1:length(zlabs$SYS_NM))
zlabs <- merge(zlabs, rps.clean, by = "SYS_NM")
zlabs <- zlabs %>%
  select(SYS_NM, order, inlet)
zlabs <- distinct(zlabs)
zlabcols <- as.data.frame(unique(zlabs$inlet))
colnames(zlabcols) <- "inlet"
zlabcols <- zlabcols %>%
  arrange(inlet)
zlabcols$colour <- c("red", "blue", "green")
zlabs <- merge(zlabcols, zlabs, by = "inlet")
zlabs <- zlabs %>%
  arrange(order)
zlabs <- zlabs$colour

elabs <- as.data.frame(dendclust.e[["labels"]][["label"]])
colnames(elabs) <- "SYS_NM"
elabs$order <- c(1:length(elabs$SYS_NM))
elabs <- merge(elabs, rps.clean, by = "SYS_NM")
elabs <- elabs %>%
  select(SYS_NM, order, inlet)
elabs <- distinct(elabs)
elabcols <- as.data.frame(unique(elabs$inlet))
colnames(elabcols) <- "inlet"
elabcols <- elabcols %>%
  arrange(inlet)
elabcols$colour <- c("red", "blue", "green")
elabs <- merge(elabcols, elabs, by = "inlet")
elabs <- elabs %>%
  arrange(order)
elabs <- elabs$colour

slabs <- as.data.frame(dendclust.s[["labels"]][["label"]])
colnames(slabs) <- "SYS_NM"
slabs$order <- c(1:length(slabs$SYS_NM))
slabs <- merge(slabs, rps.clean, by = "SYS_NM")
slabs <- slabs %>%
  select(SYS_NM, order, inlet)
slabs <- distinct(slabs)
slabcols <- as.data.frame(unique(slabs$inlet))
colnames(slabcols) <- "inlet"
slabcols <- slabcols %>%
  arrange(inlet)
slabcols$colour <- c("red", "blue", "green")
slabs <- merge(slabcols, slabs, by = "inlet")
slabs <- slabs %>%
  arrange(order)
slabs <- slabs$colour

mlabs <- as.data.frame(dendclust.m[["labels"]][["label"]])
colnames(mlabs) <- "SYS_NM"
mlabs$order <- c(1:length(mlabs$SYS_NM))
mlabs <- merge(mlabs, rps.clean, by = "SYS_NM")
mlabs <- mlabs %>%
  select(SYS_NM, order, inlet)
mlabs <- distinct(mlabs)
mlabcols <- as.data.frame(unique(mlabs$inlet))
colnames(mlabcols) <- "inlet"
mlabcols <- mlabcols %>%
  arrange(inlet)
mlabcols$colour <- c("red", "blue", "green")
mlabs <- merge(mlabcols, mlabs, by = "inlet")
mlabs <- mlabs %>%
  arrange(order)
mlabs <- mlabs$colour

rpslabs <- as.data.frame(dendclust.rps[["labels"]][["label"]])
colnames(rpslabs) <- "SYS_NM"
rpslabs$order <- c(1:length(rpslabs$SYS_NM))
rpslabs <- merge(rpslabs, rps.clean, by = "SYS_NM")
rpslabs <- rpslabs %>%
  select(SYS_NM, order, inlet)
rpslabs <- distinct(rpslabs)
rpslabcols <- as.data.frame(unique(rpslabs$inlet))
colnames(rpslabcols) <- "inlet"
rpslabcols <- rpslabcols %>%
  arrange(inlet)
rpslabcols$colour <- c("red", "blue", "green")
rpslabs <- merge(rpslabcols, rpslabs, by = "inlet")
rpslabs <- rpslabs %>%
  arrange(order)
rpslabs <- rpslabs$colour


rpsloglabs <- as.data.frame(dendclust.rpslog[["labels"]][["label"]])
colnames(rpsloglabs) <- "SYS_NM"
rpsloglabs$order <- c(1:length(rpsloglabs$SYS_NM))
rpsloglabs <- merge(rpsloglabs, rps.clean, by = "SYS_NM")
rpsloglabs <- rpsloglabs %>%
  select(SYS_NM, order, inlet)
rpsloglabs <- distinct(rpsloglabs)
rpsloglabcols <- as.data.frame(unique(rpsloglabs$inlet))
colnames(rpsloglabcols) <- "inlet"
rpsloglabcols <- rpsloglabcols %>%
  arrange(inlet)
rpsloglabcols$colour <- c("red", "blue", "green")
rpsloglabs <- merge(rpsloglabcols, rpsloglabs, by = "inlet")
rpsloglabs <- rpsloglabs %>%
  arrange(order)
rpsloglabs <- rpsloglabs$colour

```

```{r, Dendrogram plots, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 9.5, fig.width = 8.5, fig.cap = "Dendrograms by metric. Red labels are from Esperanza inlet, blue are from Nootka inlet, and green are from Tahsis inlet."}

par(oma=c(0,0,0,0))
par(mfrow=c(3,2))

clust.dist.cor.z[["labels"]] <- str_to_sentence(word(clust.dist.cor.z[["labels"]]))
dendz <- as.dendrogram(clust.dist.cor.z)
dendz %>%
    set("labels_colors", zlabs) %>%
    plot(main = "Z-scores", horiz = TRUE)

clust.dist.cor.e[["labels"]] <- str_to_sentence(word(clust.dist.cor.e[["labels"]]))
dende <- as.dendrogram(clust.dist.cor.e)
dende %>%
    set("labels_colors", elabs) %>%
    plot(main = "Log escapement", horiz = TRUE)

clust.dist.cor.s[["labels"]] <- str_to_sentence(word(clust.dist.cor.s[["labels"]]))
dends <- as.dendrogram(clust.dist.cor.s)
dends %>%
    set("labels_colors", slabs) %>%
    plot(main = "Standardized escapement", horiz = TRUE)

clust.dist.cor.m[["labels"]] <- str_to_sentence(word(clust.dist.cor.m[["labels"]]))
dendm <- as.dendrogram(clust.dist.cor.m)
dendm %>%
    set("labels_colors", mlabs) %>%
    plot(main = "Moving average", horiz = TRUE)

clust.dist.cor.rps[["labels"]] <- str_to_sentence(word(clust.dist.cor.rps[["labels"]]))
dendrps <- as.dendrogram(clust.dist.cor.rps)
dendrps %>%
    set("labels_colors", rpslabs) %>%
    plot(main = "RPS", horiz = TRUE)

clust.dist.cor.rpslog[["labels"]] <- str_to_sentence(word(clust.dist.cor.rpslog[["labels"]]))
dendrpslog <- as.dendrogram(clust.dist.cor.rpslog)
dendrpslog %>%
    set("labels_colors", rpsloglabs) %>%
    plot(main = "Log RPS", horiz = TRUE)
```

\newpage

## Tanglegrams comparing effect of metric choice on cluster analysis

```{r tanglegram z vs log, echo = FALSE, warning = FALSE, message = FALSE,  fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing the use of z-score against the use of log escapements on cluster analysis outputs."}

#png(file = "figures/tanglegram_z_v_log.png", width = 2000, height = 3000, pointsize = 36)

par(oma=c(2,2,2,2))
par(mfrow=c(1,1))

tangle1 <- dendlist(dendz, dende)
tanglegram(tangle1, lab.cex = .8, margin_inner = 3, main_left = "Z-score",
           main_right = "Log escapement", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)

#dev.off()

```
\newpage
```{r tanglegram z vs standardized esc, echo = FALSE, warning = FALSE, message = FALSE,  fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing the use of z-score against the use of standardized escapements on cluster analysis outputs."}

#png(file = "figures/tanglegram_z_v_stdesc.png", width = 2000, height = 3000, pointsize = 36)

tangle2 <- dendlist(dendz, dends)
tanglegram(tangle2, lab.cex = .8, margin_inner = 3, main_left = "Z-score",
           main_right = "Standardized escapement", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)
#dev.off()

```
\newpage
```{r tanglegram a vs m.ave, echo = FALSE, warning = FALSE, message = FALSE,  fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing the use of z-score against the use of moving average on cluster analysis outputs."}

#png(file = "figures/tanglegram_z_v_mave.png", width = 2000, height = 3000, pointsize = 36)

tangle3 <- dendlist(dendz, dendm)
tanglegram(tangle3, lab.cex = .8, margin_inner = 3, main_left = "Z-score",
           main_right = "Moving average", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)
#dev.off()

```

\newpage

```{r, Tanglegrams z vs logRPS, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 9.5, fig.width = 8.5, fig.cap = "Tanglegram of z-score vs. Log RPS"}

#png(file = "figures/a06cm_tanglegram_z_v_mave.png", width = 2000, height = 3000, pointsize = 36)

tangle4 <- dendlist(dendz, dendrpslog)
                                                                                                            
tanglegram(tangle4, lab.cex = .8, margin_inner = 3, main_left = "Z-score",
           main_right = "Log RPS", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)

#dev.off()

```

\newpage

# Analyses by pre- and post-enhancement
## Correlation plots

```{r Corr plots for z-score, pre- and post-1980, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Cross correlation plots of z-scores to compare pre- and post-enhancement."}
#prep data for correlations

cor.data.z <-escdata.filt %>%
  select(SYS_NM,Year,z) %>%
  pivot_wider(names_from="Year",values_from="z") %>%
  data.frame()

cor.data.z.pre<-cor.data.z[,1:31]
rownames(cor.data.z.pre)<-cor.data.z.pre$SYS_NM
cor.data.z.pre<-cor.data.z.pre%>%select(-SYS_NM)
cor.z.pre<-cor(t(cor.data.z.pre),use="pairwise.complete.obs")

cor.data.z.post<-cor.data.z[,c(1,32:70)]
rownames(cor.data.z.post)<-cor.data.z.post$SYS_NM
cor.data.z.post<-cor.data.z.post%>%select(-SYS_NM)
cor.z.post<-cor(t(cor.data.z.post),use="pairwise.complete.obs")

#png(file = "figures/corrplot_1980_z.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(2,1))
corrplot(cor.z.pre, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Z-scores (Pre-1980)", cex.main = .7, mar=c(0,0,1,0))
corrplot(cor.z.post, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Z-score (Post-1980)", sub = "Z-scores", cex.main = .7, mar=c(0,0,1,0))

#dev.off()
```

\newpage

```{r, Corr plots for log escapement, pre- and post-1980, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Cross correlation plots to compare log escapements pre- and post-enhancement."}

# Log escapements 

cor.data.e<-escdata.filt%>%select(SYS_NM,Year,esc.log)%>%
  pivot_wider(names_from="Year",values_from="esc.log")%>%data.frame()

cor.data.e.pre<-cor.data.e[,1:31]
rownames(cor.data.e.pre)<-cor.data.e.pre$SYS_NM
cor.data.e.pre<-cor.data.e.pre%>%select(-SYS_NM)
cor.e.pre<-cor(t(cor.data.e.pre),use="pairwise.complete.obs")

cor.data.e.post<-cor.data.e[,c(1,32:70)]
rownames(cor.data.e.post)<-cor.data.e.post$SYS_NM
cor.data.e.post<-cor.data.e.post%>%select(-SYS_NM)
cor.e.post<-cor(t(cor.data.e.post),use="pairwise.complete.obs")

#png(file = "figures/corrplot_1980_log.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(2,1))
corrplot(cor.e.pre, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Log escapement (Pre-1980)", cex.main = .7, mar=c(0,0,1,0))
corrplot(cor.e.post, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Log escapement (Post-1980)", cex.main = .7, mar=c(0,0,1,0))
#dev.off()

```

\newpage

```{r, Corr plots for standardized escapements, pre- and post-1980, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Cross correlation plots to compare standardized escapements pre- and post-enhancement."}

# Standardized escapements

cor.data.s<-escdata.filt%>%select(SYS_NM,Year,esc.stand)%>%
  pivot_wider(names_from="Year",values_from="esc.stand")%>%data.frame()

cor.data.s.pre<-cor.data.s[,1:31]
rownames(cor.data.s.pre)<-cor.data.s.pre$SYS_NM
cor.data.s.pre<-cor.data.s.pre%>%select(-SYS_NM)
cor.s.pre<-cor(t(cor.data.s.pre),use="pairwise.complete.obs")

cor.data.s.post<-cor.data.s[,c(1,32:70)]
rownames(cor.data.s.post)<-cor.data.s.post$SYS_NM
cor.data.s.post<-cor.data.s.post%>%select(-SYS_NM)
cor.s.post<-cor(t(cor.data.s.post),use="pairwise.complete.obs")

#png(file = "figures/corrplot_1980_stdesc.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(2,1))
corrplot(cor.s.pre, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Standardized escapement (Pre-1980)", cex.main = .7, mar=c(0,0,1,0))
corrplot(cor.s.post, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Standardized escapement (Post-1980)", cex.main = .7, mar=c(0,0,1,0))

#dev.off()

```

\newpage

```{r, Corr plots for moving average, pre- and post-1980, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Cross correlation plots to compare moving average pre- and post-enhancement."}

# Moving average

cor.data.m<-escdata.filt%>%select(SYS_NM,Year,m.ave)%>%
  pivot_wider(names_from="Year",values_from="m.ave")%>%data.frame()

cor.data.m.pre<-cor.data.m[,1:31]
rownames(cor.data.m.pre)<-cor.data.m.pre$SYS_NM
cor.data.m.pre<-cor.data.m.pre%>%select(-SYS_NM)
cor.m.pre<-cor(t(cor.data.m.pre),use="pairwise.complete.obs")

cor.data.m.post<-cor.data.m[,c(1,32:70)]
rownames(cor.data.m.post)<-cor.data.m.post$SYS_NM
cor.data.m.post<-cor.data.m.post%>%select(-SYS_NM)
cor.m.post<-cor(t(cor.data.m.post),use="pairwise.complete.obs")

#png(file = "figures/corrplot_1980_mave.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(2,1))
corrplot(cor.m.pre, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Moving average (Pre-1980)", cex.main = .7, mar=c(0,0,1,0))
corrplot(cor.m.post, diag = FALSE, order="hclust", method="square", tl.col="black",
         addrect = 6, tl.cex = 0.5, title = "Moving average (Post-1980)", cex.main = .7, mar=c(0,0,1,0))

#dev.off()
```

\newpage

## Tanglegrams

```{r, z-score tanglegram, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing z-scores pre- and post-enhancement (1980)"}
# Dendrogram, factored by inlet, by z-score

#png(file = "figures/tanglegram_1980_z.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(1,1))

# Pre-1980
dist.cor.z.pre <- dist(cor.z.pre)
clust.dist.cor.z.pre <- hclust(dist.cor.z.pre)
tanglelabs.z.pre <- as.data.frame(clust.dist.cor.z.pre[["labels"]])
colnames(tanglelabs.z.pre) <- "SYS_NM"
tanglelabs.z.pre <- merge(tanglelabs.z.pre, nameslookup, on = "SYS_NM")
clust.dist.cor.z.pre[["labels"]] <- tanglelabs.z.pre$shortnames
dendz.pre <- as.dendrogram(clust.dist.cor.z.pre)

# Post-1980
dist.cor.z.post <- dist(cor.z.post)
clust.dist.cor.z.post <- hclust(dist.cor.z.post)
tanglelabs.z.post <- as.data.frame(clust.dist.cor.z.post[["labels"]])
colnames(tanglelabs.z.post) <- "SYS_NM"
tanglelabs.z.post <- merge(tanglelabs.z.post, nameslookup, on = "SYS_NM")
clust.dist.cor.z.post[["labels"]] <- tanglelabs.z.post$shortnames
dendz.post <- as.dendrogram(clust.dist.cor.z.post)

# Tanglegram
tangle.z <- dendlist(dendz.pre, dendz.post)
tanglegram(tangle.z, lab.cex = .8, margin_inner = 3, main = "Z-scores", main_left = "Pre-1980",
           main_right = "Post-1980", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)

#dev.off()

```

\newpage

```{r, Log escapements tanglegram, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing log escapements pre- and post-enhancement (1980)"}

#png(file = "figures/tanglegram_1980_log.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(1,1))

# Pre-1980
dist.cor.e.pre <- dist(cor.e.pre)
clust.dist.cor.e.pre <- hclust(dist.cor.e.pre)
tanglelabs.e.pre <- as.data.frame(clust.dist.cor.e.pre[["labels"]])
colnames(tanglelabs.e.pre) <- "SYS_NM"
tanglelabs.e.pre <- merge(tanglelabs.e.pre, nameslookup, on = "SYS_NM")
clust.dist.cor.e.pre[["labels"]] <- tanglelabs.e.pre$shortnames
dende.pre <- as.dendrogram(clust.dist.cor.e.pre)

# Post-1980
dist.cor.e.post <- dist(cor.e.post)
clust.dist.cor.e.post <- hclust(dist.cor.e.post)
tanglelabs.e.post <- as.data.frame(clust.dist.cor.e.post[["labels"]])
colnames(tanglelabs.e.post) <- "SYS_NM"
tanglelabs.e.post <- merge(tanglelabs.e.post, nameslookup, on = "SYS_NM")
clust.dist.cor.e.post[["labels"]] <- tanglelabs.e.post$shortnames
dende.post <- as.dendrogram(clust.dist.cor.e.post)

# Tanglegram
tangle.e <- dendlist(dende.pre, dende.post)
tanglegram(tangle.e, lab.cex = .8, margin_inner = 3, main = "Log escapements", main_left = "Pre-1980",
           main_right = "Post-1980", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)
#dev.off()

```

\newpage

```{r, Standardized escapements tanglegram, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing log escapements pre- and post-enhancement (1980)"}

#png(file = "figures/tanglegram_1980_stdesc.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(1,1))

# Pre-1980
dist.cor.s.pre <- dist(cor.s.pre)
clust.dist.cor.s.pre <- hclust(dist.cor.s.pre)
tanglelabs.s.pre <- as.data.frame(clust.dist.cor.s.pre[["labels"]])
colnames(tanglelabs.s.pre) <- "SYS_NM"
tanglelabs.s.pre <- merge(tanglelabs.s.pre, nameslookup, on = "SYS_NM")
clust.dist.cor.s.pre[["labels"]] <- tanglelabs.s.pre$shortnames
dends.pre <- as.dendrogram(clust.dist.cor.s.pre)

# Post-1980
dist.cor.s.post <- dist(cor.s.post)
clust.dist.cor.s.post <- hclust(dist.cor.s.post)
tanglelabs.s.post <- as.data.frame(clust.dist.cor.s.post[["labels"]])
colnames(tanglelabs.s.post) <- "SYS_NM"
tanglelabs.s.post <- merge(tanglelabs.s.post, nameslookup, on = "SYS_NM")
clust.dist.cor.s.post[["labels"]] <- tanglelabs.s.post$shortnames
dends.post <- as.dendrogram(clust.dist.cor.s.post)

# Tanglegram
tangle.s <- dendlist(dends.pre, dends.post)
tanglegram(tangle.s, lab.cex = .8, margin_inner = 3, main = "Standardized escapements", main_left = "Pre-1980",
           main_right = "Post-1980", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)

#dev.off()

```

\newpage

```{r, Moving averages tanglegram, echo = FALSE, fig.align = "center", fig.height = 8, fig.width = 6, fig.cap = "Tanglegram comparing moving averages pre- and post-enhancement (1980)"}

#png(file = "figures/tanglegram_1980_mave.png", width = 2000, height = 3000, pointsize = 48)

par(mfrow=c(1,1))
# Pre-1980
dist.cor.m.pre <- dist(cor.m.pre)
clust.dist.cor.m.pre <- hclust(dist.cor.m.pre)
tanglelabs.m.pre <- as.data.frame(clust.dist.cor.m.pre[["labels"]])
colnames(tanglelabs.m.pre) <- "SYS_NM"
tanglelabs.m.pre <- merge(tanglelabs.m.pre, nameslookup, on = "SYS_NM")
clust.dist.cor.m.pre[["labels"]] <- tanglelabs.m.pre$shortnames
dendm.pre <- as.dendrogram(clust.dist.cor.m.pre)

# Post-1980
dist.cor.m.post <- dist(cor.m.post)
clust.dist.cor.m.post <- hclust(dist.cor.m.post)
tanglelabs.m.post <- as.data.frame(clust.dist.cor.m.post[["labels"]])
colnames(tanglelabs.m.post) <- "SYS_NM"
tanglelabs.m.post <- merge(tanglelabs.m.post, nameslookup, on = "SYS_NM")
clust.dist.cor.m.post[["labels"]] <- tanglelabs.m.post$shortnames
dendm.post <- as.dendrogram(clust.dist.cor.m.post)

# Tanglegram
tangle.m <- dendlist(dendm.pre, dendm.post)
tanglegram(tangle.m, lab.cex = .8, margin_inner = 3, main = "Moving average ", main_left = "Pre-1980",
           main_right = "Post-1980", cex_main = 1, highlight_distinct_edges = FALSE,
           lwd = 2, edge.lwd = 1)

#dev.off()

```

\newpage

# Statistical models

## Table of Recruits per spawner candidate models

```{r, Table of RPS candidate models, echo = FALSE, message = FALSE, warning = FALSE}

rpstab <- read_csv("data/rpsmodels.csv")

kable(rpstab, format = "latex")

#rpstab %>%
#  kbl(align = colaligns, format.args = list(big.mark = ",")) %>%
#  kable_paper(full_width = F, html_font = "Courier New") %>%
#  kable_styling(latex_options = "HOLD_position")

```

\newpage

## Table of Escapement candidate models

```{r, Table of Escapement candidate models, echo = FALSE, message = FALSE, warning = FALSE}

esctab <- read_csv("data/escapementmodels.csv")

kable(esctab, format = "latex")

#esctab %>%
#  #kbl(align = colaligns, format.args = list(big.mark = ",")) %>%
#  kable_paper(full_width = F, html_font = "Courier New") %>%
#  kable_styling(latex_options = "HOLD_position")

```

\newpage

## AIC Table for RPS candidate models

```{r, AIC Table for RPS candidate models, echo = FALSE, message = FALSE, warning = FALSE}

rpsaic <- readRDS("RDS/rpsaic.rds")

rpstab$`Degrees of freedom` <- rpsaic$df
rpstab$AIC <- rpsaic$AIC

kable(rpstab, format = "latex")

#rpstab %>%
#  kbl(align = colaligns) %>%
#  kable_paper(full_width = F, html_font = "Courier New") %>%
#  kable_styling(latex_options = "HOLD_position")

```

\newpage

## AIC Table for Escapement candidate models

```{r, AIC Table for Escapement candidate models, echo = FALSE, message = FALSE, warning = FALSE}

escaic <- readRDS("RDS/escaic.rds")

esctab$`Degrees of freedom` <- escaic$df
esctab$AIC <- escaic$AIC

kable(esctab, format = "latex")

#esctab %>%
#  kbl(align = colaligns) %>%
 # kable_paper(full_width = F, html_font = "Courier New") %>%
  #kable_styling(latex_options = "HOLD_position")

```

\newpage

## Effects plot of log RPS by year

```{r, Effects plot log rps by year, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 5, fig.width = 8, fig.cap = "Effects plots of Recruits per spawner by year"}

par(mfrow=c(1,1))

EPC_RPS_YEAR <- readRDS("RDS/EPC_RPS_YEAR.rds")
plot(EPC_RPS_YEAR)

```

\newpage

## Effects plot of log RPS by total releases

```{r, Effects plots log rps by totrel, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 5, fig.width = 8, fig.cap = "Effects plots of Recruits per spawner by total releases"}

par(mfrow=c(1,1))

EPC_RPS_TOTREL <- readRDS("RDS/EPC_RPS_TOTREL.rds")
plot(EPC_RPS_TOTREL)
```

\newpage

## Effects plot of escapement by correlation coefficient

```{r, Effects plots escapement x corcoef, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 5, fig.width = 8, fig.cap = "Effects plots of Escapement by correlation coefficient"}

EP_ESC_COR <- readRDS("RDS/EP_ESC_COR.rds")
plot(EP_ESC_COR)

```

\newpage

## Effects plot of escapement by distance from enhancement

```{r, Effects plots escapement x dist, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 5, fig.width = 8, fig.cap = "Effects plots of Escapement by distance from enhancement"}

EP_ESC_DIST <- readRDS("RDS/EP_ESC_DIST.rds")
plot(EP_ESC_DIST)

```

\newpage

## Effects plot of escapement by total hatchery releases

```{r, Effects plots escapement esc x totrel, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", fig.height = 5, fig.width = 8, fig.cap = "Effects plots of Escapement by correlation coefficient"}

EP_ESC_TOTREL <- readRDS("RDS/EP_ESC_TOTREL.rds")
plot(EP_ESC_TOTREL)

```
